---
title: "WTAP/Viriliser RNAseq Analysis"
author: "Matthew Parker"
date: "19 June 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## WTAP/Viriliser RNAseq Analysis

```{r}
library(tximport)
library(edgeR)
library(biomaRt)

ensembl = useMart(biomart="ENSEMBL_MART_ENSEMBL", 
                  host="grch37.ensembl.org",
                  path="/biomart/martservice",
                  dataset="hsapiens_gene_ensembl")
annot <- getBM(attributes=c('ensembl_transcript_id',
                            'ensembl_gene_id',
                            'hgnc_symbol',
                            'description'),
               mart=ensembl)
tx2gene <- annot[,1:2]
write.table(tx2gene, '~/Documents/analysis_notebooks/m6a/wtap_viril/zenodo/output/transcript2gene.tsv', sep='\t', quote=FALSE)
annot <- annot[!duplicated(annot$ensembl_gene_id),]
annot <- annot[,c(2:4)]
```

```{r}
fn = Sys.glob('~/Documents/analysis_notebooks/m6a/wtap_viril/zenodo/salmon_counts/*hg19_incl_ncrna_salmon_output/quant.sf')
print("## File names:")
print(paste(fn, sep="\n"))
sample.names = unlist(strsplit(basename(dirname(fn)), '_', 2))[((1:12)*8 - 7)]
print("## Sample names:")
print(paste(sample.names, sep='\n'))
names(fn) = sample.names
salmon.gene = tximport(fn, type='salmon', tx2gene=tx2gene, dropInfReps=TRUE)
counts = salmon.gene$counts
len = salmon.gene$length
```

##Preparing data for edgeR

taken from tximport instructions for converting salmon counts to edgeR compatible format

```{r}
normMat <- len/exp(rowMeans(log(len)))
o <- log(calcNormFactors(counts/normMat)) + log(colSums(counts/normMat))
y <- DGEList(counts)
y$offset <- t(t(log(normMat)) + o)
keep <- rowSums(cpm(y) > 1) >= 2
y <- y[keep, , keep.lib.sizes=FALSE]
plotMDS(y, method='bcv', top=200)
# y is now ready for estimate dispersion functions see edgeR User's Guide
```

Build design matrix:

```{r}
groups <- factor(substr(sample.names, 1, nchar(sample.names) - 1))
design <- model.matrix(~0 + groups)
rownames(design) <- sample.names
colnames(design) <- levels(groups)
design
```

estimate dispersions with edgeR using design matrix, and fit model

```{r}
y <- estimateDisp(y, design=design)
fit <- glmFit(y, design)
```

Check KD v control to see if WTAP and viriliser are knocked down

```{r}
kd_vs_cntrl_cyto <- glmLRT(fit, contrast=makeContrasts('VWC - CC', levels=design))
plotMD(kd_vs_cntrl_cyto)

topTagsAnnot <- function(fitObj) {
  res <- merge(topTags(fitObj, n=Inf),
               annot,
               by.x=0,
               by.y='ensembl_gene_id',
               all.x=TRUE,
               all.y=FALSE)
  rownames(res) <- res$Row.names
  res <- res[,c(7,2:6,8)]
  res <- res[order(-abs(res$logFC)),]
  return (res)
}

kd_vs_cntrl_cyto.res <- topTagsAnnot(kd_vs_cntrl_cyto)
# KIAA1429 aka Virilizer
kd_vs_cntrl_cyto.res['ENSG00000164944',]
```

```{r}
# WTAP
kd_vs_cntrl_cyto.res['ENSG00000146457',]
```

Now we can look at differences between the nuclear cyto ratio in VW v Cntrl

```{r}
kd_vs_cntrl_nc_ratio <- glmLRT(fit,
                               contrast=makeContrasts('(VWN - VWC) - (CN - CC)',
                                                      levels=design))
kd_vs_cntrl_nc_ratio.res <- topTagsAnnot(kd_vs_cntrl_nc_ratio)
kd_vs_cntrl_nc_ratio.res <- kd_vs_cntrl_nc_ratio.res[kd_vs_cntrl_nc_ratio.res$logCPM > 2,]
head(kd_vs_cntrl_nc_ratio.res, n=20)
```




```{r}
plotMD(kd_vs_cntrl_nc_ratio)
```

```{r}
nc <- as.data.frame(cpm(y, normalized.lib.sizes = TRUE, log=TRUE))
all_res <- merge(kd_vs_cntrl_nc_ratio.res, nc, by=0, all.x=TRUE, all.y = FALSE)
rownames(all_res) <- all_res$gene_id
all_res <- all_res[,3:ncol(all_res)-1]
write.table(all_res, 
            file='~/Documents/analysis_notebooks/m6a/wtap_viril/zenodo/output/nuclear_accumulation_results.tsv',
            sep='\t',
            quote=FALSE)
```
